### 2018-04-20 RefereeReportResponse-17Apr2018.txt
Start addressing referee comments

----------------------------------------------------------------------
Report of Referee A -- LQ16046/Thyagarajan
----------------------------------------------------------------------

In this paper, the authors propose using the bispectrum phase (i.e.,
closure phase in radio interferometry) as a way to detect the 21cm
signal from cosmic reionization. This is attractive because
antenna-based gain factors naturally cancel out of the observable
quantities, so a reported detection of the cosmological signal would
be more robust to systematics. Given the difficulties in calibrating
low-frequency interferometers to high precision, I think this work is
extremely important. I would be happy to recommend acceptance after my
comments below are addressed. This paper is suitable for publication
in PRL rather than (say) PRD because for years the 21cm community has
searched for an alternative to the power spectrum. The results are
either not scientifically interesting or are not observationally
realistic. This paper is a rare example that does not suffer from
these shortcomings.

Specific comments:

1) Introduction: the comment about the photon-to-baryon ratio is a
rather strange comment, because the statement is only true if one
ignores the CMB in calculating the ratio.

[Response] Modified to "photon-to-baryon ratio rose significantly"

2) Penultimate paragraph of introduction: I think it's debatable
whether current experiments have already "obtained sufficient
observational data"

[Response] Modified "sufficient" to "significant"

3) The summation/product notation should be revised to conform to
convention. Generally, if one writes something like \sum_{ij}, I
interpret this to mean the double sum \sum_{i=1}^N \sum_{j=1}^N. With
N = 3 antennas in this paper, I would thus interpret the sums to give
rise to 9 terms, including the i = j terms. Instead, I think the
authors mean a single sum over 1-2, 2-3, 3-1.

[Response] The referee is correct. We mean a single sum over 1-2, 2-3,
and 3-1.

4) I am not sure why the equation for B_{\Del}^m does not contain more
terms. Each V_ij has the form (signal + noise), so when I multiply
three such quantities together I should get terms that look like
signal^2 noise and signal noise^2. The signal^2 noise term has zero
mean if the noise on a visibility has zero mean, while the signal
noise^2 term will have zero mean if the different baselines have
independent noise. So I could believe that they're zero mean, but then
so is the pure noise term in the equation, and so I don't understand
why only that pure noise term was included.

[Response] Due to constraints on space, we include all cross-terms and
noise-only terms in B_\nabla^N. Thus it comprises of all terms including
cross-terms that contain visibility noise. And in our analysis, the
triads are made of antennas i \ne j \ne k. Thus B_\nabla^N is zero-mean.

5) On a different, but related note, I think that the noise term
B_{\Del}^N only has mean zero if one assumes that the noise
contributions from different baselines are uncorrelated. It's not
sufficient to just say that it has zero mean because it contains V^N.

[Response] The referee is correct. We assume the noise across baselines
is uncorrelated.

6) Ultimately, I felt that the authors should probably have discussed
a little more whether in practice one would expect a) negligible gain
factors that cannot be parameterized on a per-antenna basis, and b)
noise that is necessarily independent between baselines.

[Response] In practice, both these items can only be evaluated from
data. Since these non-ideal effects are instrument specific, and
difficult to estimate without data, we have left it for the data
analysis paper. We will present the results of such data analysis from
HERA in a forthcoming paper. For purposes of concept demonstration, in
this paper we have assumed non-antenna based gains are negligible, and
noise is uncorrelated across baselines.

7) I think the authors should clarify that \delta \phi_{ij}^N is *not*
the phase of B_{\Del}^N, lest the unsuspecting reader think otherwise.
My understanding is that taking the natural logarithm of a complex
number to get at the phase, we have something like ln(s+n) = ln s + ln
(1 + n / s) ~ ln s + n/s. So the phase \delta \phi_{ij}^N is
noise-to-signal ratio itself, rather than the phase of the noise. Is
that correct?

[Response] That is correct. In fact, we have stated the same fact in
the first sentence of last paragraph of section II. The referee's
point about the unsuspecting reader is well taken. The noise in the
visibility measurement translates to phase fluctuations (relative to
true sky phase) whose magnitude (or RMS) is a measure of 1/(S/N).
The phase fluctuations caused by noise are NOT the phase of the noise.
It also depends on the sky signal strength. It is only a notation we
have employed, but it has to be interpreted as not the phase of the
noise itself but fluctuations to phase caused by presence of noise
and these fluctuations \delta\phi_{ij}^N are perturbations over the
true phase and have an RMS 1/(S/N). We have modified text below
Equation 1 to clarify this point.

8) In the final paragraph of Section II, it seems like the authors are
ignoring variance from the foregrounds. Some foregrounds (e.g.
extremely bright point sources) would be inappropriate to describe in
terms of a variance, but for others (e.g., dim, unresolved point
sources), a statistical distribution like a Poisson distribution would
be an appropriate description. That would certainly contribute to the
variance.

9) Note that the previous point also affects comment 7) above. If
\delta \phi_{ij}^N goes like n/s, then yes, if n is Gaussian and s is
some definite signal with no variance, then n/s will be Gaussian. But
if s has its own distribution, then n/s may not be Gaussian. I'm not
asking for a full treatment of all probability distributions, but
(assuming that I'm right about these points), some discussions of the
caveats would be important.

10) In the 3rd paragraph of Section IV, shouldn't it be |V_F| rather
than V_F? At least that's what I would've thought given the definition
of rho.

[Response] Correct. Fixed now.

11) When the authors define the cross-power spectrum, it would be
worth mentioning which sorts of axes they imagine computing these
cross spectra.

[Response] This is discussed in Section IV B. Cross-power spectrum
would be used in cases where redundancy cannot be assumed. In section
IV B, incoherent averaging is done over corss-power spectrum. For
example, the sky is redundant in LST, that is it repeats every day at
a given LST. In this case, the closure phases themselves can be averaged
coherently before computing the power spectrum. Similarly, if the
baselines are truly redundant, those closure phases can be coherently
averaged. But if such a redundancy in baselines or neighboring LST scans
cannot be assumed, one would have to perform incoherent averaging on
cross-power spectra along axes which lack redundancy. The assumption
behind incoherent averaging of cross-power spectra is that the
cosmological signal is statistically isotropic in space whereas the
foregrounds and noise measurements can be averaged down (though not
as fast as coherent averaging).

12) I am curious: does Equation 3 vanish (in expectation) if I try to
form a cross-power spectrum with two triads that share one edge?

[Response] If the two triads are not redundant but share an edge, it
depends on whether the measurement is sky noise dominated or receiver
noise dominated. The former is generally the case at low frequencies.
In such a scenario, there will be a non-zero correlation between the
measurements on the two triads and thus the expectation of equation 3
will not necessarily vanish. Even when the measurements are receiver
noise dominated, and sufficient coherent averaging has been performed
to an extent that receiver noise contributions have been rendered
negligible, the correlations from residual sky noise may then be
visible between the two triads.

13) In Section IV C, "HERA-47" is not defined. I assume this is HERA
with 47 dishes, but how are they arranged in this intermediate stage
of HERA's buildout?

[Response] The HERA-47 layout has been illustrated in the companion
paper [Carilli et al. 2018]. We are unable to describe further due to
lack of space but have modified appropriately with a citation to the
companion paper.

14) Figure 3 needs to be made considerably easier to read. For one, it
is the main result of the paper and deserves to be a much bigger plot.
(This is not because I'm old and have bad eyesight...let's just say
that I am within 10 years of the age of the lead author of this
paper). Some of the colors are quite similar, and I would also like to
be able to know what the different curves are plotting without having
to read the caption.

15) The signal-to-noise ratio (inferred from Figure 3) seems to be
quite high even at high k. Does this trend continue beyond the values
plotted? If so, this is a very interesting property of the bispectrum
phase, because the regular power spectrum loses signal to noise
quickly with k. I wonder if this is a promising way to get at some
high k information? I guess it's hard to tell because it's not clear
what the intuition is like in terms of the scattering of power over
different k's.

16) Do the conclusions of the last paragraph of Section II require
assuming that the HI fluctuations are Gaussian? (Which of course they
aren't, although I'm fine with that assumption being made for
simplicity for this paper).

----------------------------------------------------------------------
Report of Referee B -- LQ16046/Thyagarajan
----------------------------------------------------------------------

I feel this paper should be published in PRL. It proposes an elegant
solution to a challenge that has plagued interferometers attempting to
exploit the faint radio signal from neutral hydrogen gas at high
redshifts as a cosmological tool for understanding the epoch of
reionization.

Indeed, my main suggestion is to expand the scope of the paper by
mentioning in the abstract and introduction that the technique has
broader application than just to measurements of the EoR. I believe
this broader focus would increase the impact of the paper
significantly. Cosmological measurements of HI at lower redshifts (<
2) are also being pursued vigorously by a variety of radio
interferometers (CHIME, Tianlai, HIRAX) that face the same challenges
as EoR instruments such as HERA. The simulations described in the
paper using HERA as an example are equally relevant for these other
interferometers, which also must precisely calibrate instrumental
gains in the presence of strong astrophysical foregrounds. There is
considerable interest in the cosmological community to develop HI
intensity mapping as an efficient tool for mapping large-scale
structure. The hoped-for payoff is improved precision on cosmological
parameters such as the neutrino mass and the equation of state of the
dark energy. Outside of the introduction, I believe that little or
nothing would need to be changed in the rest of the paper.

I do have one minor suggestion:

Just after equation 1, a phrase should be reversed to read “small
perturbations due to EoR HI signal and noise, respectively”.

[Response] Fixed.

### 2018-04-23 RefereeReportResponse-17Apr2018.txt
Update on Overleaf.

@@
 [Response] Due to constraints on space, we include all cross-terms and
-noise-only terms in B_\nabla^N. Thus it comprises of all terms including
-cross-terms that contain visibility noise. And in our analysis, the
-triads are made of antennas i \ne j \ne k. Thus B_\nabla^N is zero-mean.
+noise-only terms under B_\nabla^N. Thus it comprises of all terms 
+including cross-terms that contain visibility noise. And in our
+analysis, the triads are made of antennas i \ne j \ne k. Thus
+B_\nabla^N is zero-mean.
 
@@
 
-[Response] The referee is correct. We assume the noise across baselines
-is uncorrelated.
+[Response] The referee is correct. We assume the noise across
+baselines is uncorrelated.
 
@@
 this paper we have assumed non-antenna based gains are negligible, and
-noise is uncorrelated across baselines.
+noise is uncorrelated across baselines. Diagnostics from some
+preliminary data is discussed in a companion paper Carilli et al.
+(2018). Our forthcoming paper will include a full analysis detailing
+the propagation of such non-ideal beahvior into the final power
+spectrum.
 
@@
 
+[Response] In this analysis, we have taken one realization of the sky
+as available from low frequency surveys, primarily the Haslam map at
+408 MHz for the diffuse sky. Thus the sky model has no variance across
+repeating LST or redundant triads because a single realization and
+a fully deterministic sky is used, unlike noise.
+
 9) Note that the previous point also affects comment 7) above. If
@@
 caveats would be important.
+
+[Response] The referee would be right if we using statistical
+realizations of the sky model. However, since we are using a fully
+deterministic sky model, the effects of sky distribution will not
+apply.
 
@@
 
-[Response] This is discussed in Section IV B. Cross-power spectrum
+[Response] Section IV B addresses this question. Cross-power spectrum
 would be used in cases where redundancy cannot be assumed. In section
-IV B, incoherent averaging is done over corss-power spectrum. For
+IV B, incoherent averaging is done over cross-power spectrum. For
 example, the sky is redundant in LST, that is it repeats every day at
-a given LST. In this case, the closure phases themselves can be averaged
-coherently before computing the power spectrum. Similarly, if the
-baselines are truly redundant, those closure phases can be coherently
-averaged. But if such a redundancy in baselines or neighboring LST scans
-cannot be assumed, one would have to perform incoherent averaging on
-cross-power spectra along axes which lack redundancy. The assumption
-behind incoherent averaging of cross-power spectra is that the
-cosmological signal is statistically isotropic in space whereas the
-foregrounds and noise measurements can be averaged down (though not
-as fast as coherent averaging).
+a given LST. In this case, the closure phases themselves can be
+averaged coherently before computing the power spectrum. Similarly, if
+the baselines are truly redundant, those closure phases can be
+coherently averaged. But if such a redundancy in baselines or
+neighboring LST scans cannot be assumed, one would have to perform
+incoherent averaging on cross-power spectra along these axes which
+lack redundancy. The assumption behind incoherent averaging of
+cross-power spectra is that the cosmological signal is statistically
+isotropic in space whereas the foregrounds and noise measurements vary
+and can be averaged down (though not as fast as coherent averaging).
 
@@
 
-[Response] If the two triads are not redundant but share an edge, it
-depends on whether the measurement is sky noise dominated or receiver
-noise dominated. The former is generally the case at low frequencies.
-In such a scenario, there will be a non-zero correlation between the
-measurements on the two triads and thus the expectation of equation 3
-will not necessarily vanish. Even when the measurements are receiver
-noise dominated, and sufficient coherent averaging has been performed
-to an extent that receiver noise contributions have been rendered
-negligible, the correlations from residual sky noise may then be
-visible between the two triads.
+[Response] If the two triads are not redundant but share an edge (two
+antennas in common), it depends on whether the measurement is sky
+noise dominated or receiver noise dominated. The former is generally
+the case at low frequencies. In such a scenario, there will be a
+non-zero correlation between the measurements on the two triads and
+thus the expectation of equation 3 will not necessarily vanish. Even
+when the measurements are receiver noise dominated, and sufficient
+averaging has been performed to an extent that receiver noise
+contributions have been rendered negligible, the correlations from
+residual sky noise may then be visible between the two triads.
 
@@
 
-[Response] The HERA-47 layout has been illustrated in the companion
-paper [Carilli et al. 2018]. We are unable to describe further due to
-lack of space but have modified appropriately with a citation to the
-companion paper.
+[Response] We have modified appropriately with a citation to the
+companion paper [Carilli et al. 2018], where the HERA-47 layout has
+been illustrated in the companion paper. We are unable to describe
+further due to lack of space.
 
@@
 to read the caption.
+
+[Response] The referee's concern is appreciated. We have modified the
+figure and employed colors with greater contrast. We have also
+provided more descriptive legends. Some of the redudnant axes labels
+have been suppressed in favor of more space for the figure's contents.
 
@@
 simplicity for this paper).
+
+[Response] If the referee is referring to HI fluctuations in
+brightness temperature (which typically aren't Gaussian), no such
+assumption is made or necessary. If the referee is referring to
+perturbations in bi-spectrum phase caused by HI fluctuations, again a
+Gaussian assumption for those fluctuations is neither made nor
+necessary. Such an assumption is only made for the noise fluctuations
+for an easier treatment of its behavior with averaging.
 
@@
 
+[Response] Added a sentence in the Abstract, and a paragraph in the
+Introduction about applicability to other 21 cm intensity mapping
+experiments that will measure the BAO.
+
 I do have one minor suggestion:
@@
 
-[Response] Fixed.
+[Response] Corrected.

### 2018-04-25 RefereeReportResponse-17Apr2018.txt
Update on Overleaf.

@@
 [Response] The referee is correct. We mean a single sum over 1-2, 2-3,
-and 3-1.
+and 3-1. A sentence has been added above equation 1 to clarify the notation we have adopted.
 
@@
 analysis, the triads are made of antennas i \ne j \ne k. Thus
-B_\nabla^N is zero-mean.
+B_\nabla^N is zero-mean (we assume baselines have independent noise, 
+see below).
 
@@
 (2018). Our forthcoming paper will include a full analysis detailing
-the propagation of such non-ideal beahvior into the final power
+the propagation of such non-ideal behavior into the final power
 spectrum.
@@
 
-[Response] In this analysis, we have taken one realization of the sky
-as available from low frequency surveys, primarily the Haslam map at
-408 MHz for the diffuse sky. Thus the sky model has no variance across
-repeating LST or redundant triads because a single realization and
-a fully deterministic sky is used, unlike noise.
+[Response] In this analysis, the ensemble considered is for each 
+frequency channel and not over frequency channels since the frequency 
+structure is to be preserved. Such an ensemble for noise could be 
+obtained across (redundant) triads and repeating LST on different days. However, a single realization of the foregrounds (whether point sources 
+or diffuse emission) coming from low frequency radio surveys such as 
+NVSS, Haslam map at 408 MHZ does not vary across redundant triads and repeating LST.
+Thus the sky model has no variance because a single realization (in other words, a fully deterministic sky) is used,
+unlike noise.
 
@@
 deterministic sky model, the effects of sky distribution will not
-apply.
+be applicable.
 
@@
 coherently averaged. But if such a redundancy in baselines or
-neighboring LST scans cannot be assumed, one would have to perform
-incoherent averaging on cross-power spectra along these axes which
-lack redundancy. The assumption behind incoherent averaging of
-cross-power spectra is that the cosmological signal is statistically
-isotropic in space whereas the foregrounds and noise measurements vary
-and can be averaged down (though not as fast as coherent averaging).
+time cannot be assumed, one would have to perform incoherent averaging 
+on cross-power spectra along these axes which lack redundancy. The assumption behind incoherent averaging of cross-power spectra is that 
+the cosmological signal is statistically isotropic in space whereas 
+the foregrounds and noise measurements vary and can be averaged down (though not as fast as coherent averaging).
 
@@
 companion paper [Carilli et al. 2018], where the HERA-47 layout has
-been illustrated in the companion paper. We are unable to describe
-further due to lack of space.
+been illustrated.
 
@@
 figure and employed colors with greater contrast. We have also
-provided more descriptive legends. Some of the redudnant axes labels
+provided more descriptive legends. Some of the redundant axes labels
 have been suppressed in favor of more space for the figure's contents.
@@
 different k's.
+
+[Response] This is a very interesting point that we have also been discussing. We do not have a conclusive intuition yet. This is 
+being actively investigated in a follow up work and will address this question in a forthcoming paper. It will involve evaluating the 
+k-space response of bi-spectrum phase from a wide range of EoR power spectrum models and directly compare with expectations from a 
+power spectrum approach.
 

### 2018-04-26 ms_changes.tex
Include ms_changes.tex in the repo, update ms.tex

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[
reprint,
superscriptaddress,
% groupedaddress,
amsmath,
amssymb,
aps,
prd
% prl,
% showpacs,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{amssymb}   % for math
\usepackage{hyperref}% add hypertext capabilities
\usepackage{aas_macros}
\usepackage[caption=false]{subfig}
\usepackage[final]{changes}
\definechangesauthor[name={Nithyanandan Thyagarajan}, color=red]{NT}
\definechangesauthor[name={Chris Carilli}, color=blue]{CC}
\definechangesauthor[name={Bojan Nikolic}, color=green]{BN}
\setremarkmarkup{(#2)}
% \usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

\newcommand{\HI}{H\,{\sc i}}

\begin{document}

\preprint{APS/123-QED}

\title{Detecting Cosmic Reionization using Bi-Spectrum Phase}% Force line breaks with \\
% \thanks{A footnote to the article title}%

\author{Nithyanandan Thyagarajan}
\email{t\_nithyanandan@nrao.edu}
\homepage{https://tnithyanandan.wordpress.com/}
\thanks{Nithyanandan Thyagarajan is a Jansky Fellow of the National Radio Astronomy Observatory.}
\affiliation{National Radio Astronomy Observatory, Socorro, NM 87801, USA}
\affiliation{Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA}
% \altaffiliation[Also at ]{Physics Department, XYZ University.}%Lines break automatically or can be forced with \\
\author{Chris L. Carilli}%
% \email{Second.Author@institution.edu}
\affiliation{National Radio Astronomy Observatory, Socorro, NM 87801, USA}
\affiliation{Astrophysics Group, Cavendish Laboratory, University of Cambridge, Cambridge CB3 0HE, UK}

% \collaboration{MUSO Collaboration}%\noaffiliation

\author{Bojan Nikolic}
% \homepage{http://www.Second.institution.edu/~Charlie.Author}
\affiliation{Astrophysics Group, Cavendish Laboratory, University of Cambridge, Cambridge CB3 0HE, UK}%
% \affiliation{
%  Third institution, the second for Charlie Author
% }%

% \collaboration{CLEO Collaboration}%\noaffiliation

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract} 
  Detecting neutral Hydrogen (H\,{\sc i}) via the 21~cm line emission from the intergalactic medium at $z\gtrsim 6$ has been identified as one of the most promising probes of the epoch of cosmic reionization -- a major phase transition of the Universe. However, these studies face severe challenges imposed by the bright foreground emission from cosmic objects. Current techniques require precise instrumental calibration to separate the weak H\,{\sc i} line signal from the foreground continuum emission. We propose to mitigate this calibration requirement by using \replaced[id=NT]{measurements of the interferometric bi-spectrum phase}{the phase of the bi-spectrum from interferometric measurements}. Bi-spectrum phase is unaffected by antenna-based direction-independent calibration errors and hence for a compact array it depends on the sky brightness distribution only (subject to the usual thermal-like noise). We show that the bi-spectrum phase of foreground synchrotron continuum has a characteristically smooth spectrum relative to the cosmological line signal. The two can be separated effectively by exploiting this spectral difference using Fourier techniques, while eliminating the need for precise antenna-based calibration of phases introduced by the instrument, and the ionosphere, inherent in existing approaches. Using fiducial models for continuum foregrounds, and for the cosmological H\,{\sc i} signal, we show the latter should be detectable in bi-spectrum phase spectra, with reasonable significance at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$, using existing instruments. \added[id=NT]{Our approach will also benefit other H\,{\sc i} intensity mapping experiments that face similar challenges, such as those measuring Baryon Acoustic Oscillations (BAO).}
\end{abstract}

% \begin{abstract}
% An article usually includes an abstract, a concise summary of the work
% covered at length in the main body of the article. 
% \begin{description}
% \item[Usage]
% Secondary publications and information retrieval purposes.
% \item[PACS numbers]
% May be entered using the \verb+\pacs{#1}+ command.
% \item[Structure]
% You may use the \texttt{description} environment to structure your abstract;
% use the optional argument of the \verb+\item+ command to give the category of each item. 
% \end{description}
% \end{abstract}

\pacs{}% PACS, the Physics and Astronomy
                             % Classification Scheme.
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired

% \pacs{Valid PACS appear here}% PACS, the Physics and Astronomy
%                              % Classification Scheme.
% %\keywords{Suggested keywords}%Use showkeys class option if keyword
%                               %display desired

\maketitle

%\tableofcontents

\section{Introduction}\label{sec:intro}

The hydrogen gas that dominates the baryon content of the Universe has undergone two phase transitions over cosmic history. The first was the transition from fully ionized to fully neutral gas during cosmic \emph{recombination\/} $\sim 300,000$ years after the Big Bang ($z\approx 1100$). This epoch has been quantified in exquisite detail through cosmic microwave background radiation studies \cite{planck15i}. The second transition is known as cosmic \emph{reionization}, when light from the first stars and black holes reionized the neutral Hydrogen (\HI) gas that pervaded the early Universe. Current indirect constraints suggest this second transition occurred a few hundred Myr to $\sim 1$~Gyr ($z\gtrsim 6$) \cite{gre17a} after the Big Bang. During this epoch, the photon-to-baryon ratio \replaced[id=NT]{rose significantly}{eventually exceeded unity}, reionizing the entire Universe except gas bound to galaxies. The astrophysical processes during this epoch shaped the evolution of galaxies and large-scale structure. However, unlike recombination, the timing and process of cosmic reionization remain poorly understood. The redshifted 21~cm spectral line from the electron spin-flip transition in \HI\ atom, has been recognized as the most promising tool to unravel the physical processes involved in cosmic reionization, and the evolution of large scale structure during the formation of the first galaxies \cite{sun72,sco90,mad97,toz00,ili02,fan02,fan06,bar07,mor10}.

Our Galaxy and intervening radio galaxies emit synchrotron radiation in the same frequency band ($\sim$~100--200~MHz) as the redshifted \HI\ signal from the epoch of reionization (EoR), with a brightness temperature roughly $\gtrsim 10^4$ times higher than the EoR \HI\ signal. However, the synchrotron foregrounds have a smooth spectrum whereas the EoR \HI\ signal will appear as small fluctuations superimposed on the smooth foreground spectrum. The hope of separating the cosmic line signal from the Galactic and extragalactic foregrounds has spawned a number of low-frequency instruments (eg. \cite{par10,tin13,van13}), with sensitivities sufficient for a statistical detection of the EoR \HI\ signal using power spectrum methodology \cite{thy13,bea13}.  However, despite obtaining \replaced[id=NT]{significant}{sufficient} observational data, these experiments remain dynamic range limited due to the coupling of the strong foreground emission to the EoR \HI\ signal via instrumental effects, particularly systematics related to precision of calibration, and the intrinsic chromatic instrumental response \cite{thy15a,thy15b,thy16,dat09,dat10,tro16, pac13,ali15,patil17,pob15,liu10,zhe14,barry16,sie17,dil17}.

We present an alternate approach using the phase of the complex bi-spectrum (also referred to in radio interferometry as closure phase) to detect the signal from the EoR. This quantity is impervious to antenna-based complex gains introduced by the instrument and the ionosphere, and thus can remove the need for high-precision calibration currently impeding existing approaches\cite{car16}.

\added[id=NT]{Further, \HI\ intensity mapping experiments, such as CHIME \cite{ban14}, HIRAX \cite{new16}, and Tianlai \cite{xu15} will map \HI\ on large-scales at $1\lesssim z \lesssim 3$ tracing the matter density fluctuations, and thus measure the Baryon Acoustic Oscillations (BAO) that can be used as a standard ruler in constraining the {\it Dark Energy} equation of state. The challenges in these experiments are expected to be of a similar magnitude to that in 21~cm EoR experiments, and thus can potentially benefit from our approach.}

\section{Properties of Bi-spectrum Phase}\label{sec:CPinfo}

The bi-spectrum in the context of interferometry has been investigated in \cite{jen58,kul89,tay99,tho01,mon06} and recently revisited in \cite{car18}. Consider a triad of antennas, $\{a, b, c\}$, indexed by $i$. $V_{ij}^\textrm{m}(f)$ is the spatial coherence measured at frequency, $f$, between antenna pairs, $\{ab, bc, ca\}$, indexed by $ij$. An interferometer measures $V_{ij}^\textrm{m}(f) = g_i(f)\, g_j^*(f)\, V_{ij}^\textrm{T}(f) + V_{ij}^\textrm{N}(f)$, the  true spatial coherence, $V_{ij}^\textrm{T}(f)$, corrupted by multiplicative antenna gains and ionospheric distortions, $g_i$, and additive thermal-like noise, $V_{ij}^\textrm{N}(f)$. \replaced[id=NT]{We}{It is pertinent to} express $V_{ij}^\textrm{T}(f) = V_{ij}^\textrm{F}(f) + V_{ij}^\textrm{H{\sc i}}(f)$, where, $V_{ij}^\textrm{F}(f)$ and $V_{ij}^\textrm{H{\sc i}}(f)$ are \replaced[id=NT]{due to}{contributions from} foreground synchrotron and EoR \HI\ signal, respectively. The measured bi-spectrum is $B_\nabla^\textrm{m}=\prod_{ij} V_{ij}^\textrm{m}=B_\nabla^\textrm{T}\prod_i |g_i|^2+B_\nabla^\textrm{N}$, where, the dependence on $f$ has been omitted for convenience (hereafter so, unless indicated). \added[id=NT]{In our notation, $ij$ only indexes three specific cyclic antenna pairs $\{ab, bc, ca\}$ and not all possible combinations of $i$ and $j$.} \deleted[id=NT]{The noise term,} $B_\nabla^\textrm{N}$, \added[id=NT]{encompassing all terms containing $V_{ij}^\textrm{N}$ including cross-terms,} has zero mean since it contains $V_{ij}^\textrm{N}$, \added[id=NT]{which is assumed to be uncorrelated between antenna pairs}. The phase of $B_\nabla^\textrm{m}$ is:
\begin{align}
  \phi_\nabla^\textrm{m} &= \sum_{ij}\phi_{ij}^\textrm{m} = \phi_\nabla^\textrm{T} + \delta\phi_\nabla^\textrm{N} = \left(\phi_\nabla^\textrm{F} + \delta\phi_\nabla^\textrm{H{\sc i}}\right) + \delta\phi_\nabla^\textrm{N} \label{eqn:cpphase-sum},
\end{align}
where, $\phi_{ij}^\textrm{m} = \phi_i - \phi_j + \phi_{ij}^\textrm{T} + \delta\phi_{ij}^\textrm{N}$, and $\phi_i$ are the phases of $V_{ij}^\textrm{m}$ and $g_i$, respectively. $\delta\phi_\nabla^\textrm{H{\sc i}}$ and $\delta\phi_\nabla^\textrm{N}$ \added[id=NT]{are not just determined by the phases of $V_{ij}^\textrm{H{\sc i}}$ and $V_{ij}^\textrm{N}$ but are small} \deleted[id=NT]{may be considered as small} perturbations due to \replaced[id=NT]{the EoR \HI\ signal and noise}{noise and EoR \HI\ signal} respectively, superimposed on dominant phase, $\phi_\nabla^\textrm{F}$, due to foregrounds, \added[id=NT]{and depend on $\rho_{ij}^\textrm{H{\sc i}}\equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{H{\sc i}}|$ and $\rho_{ij}^\textrm{N}\equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{N}|$ respectively.} \replaced[id=NT]{The small-perturbation formalism}{and} is valid when \replaced[id=NT]{$\rho_{ij}^\textrm{H{\sc i}}\gg 1$ and $\rho_{ij}^\textrm{N}\gg 1$}{$\rho_{ij}^\textrm{N} \equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{N}| \gg 1$ and $\rho_{ij}^\textrm{H{\sc i}} \equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{H{\sc i}}| \gg 1$}, as is usually the case in \replaced[id=NT]{low-frequency cosmology}{EoR \HI\ } experiments. $\phi_\nabla^\textrm{m}$ is independent of the antenna gains and measures $\phi_\nabla^\textrm{T}$, corrupted only by noise, $\delta\phi_\nabla^\textrm{N}$. Hence, $\phi_\nabla^\textrm{m}$ contains information purely about the sky's spatial structure, with certain geometric properties \cite[see][]{mon07}.

$\delta\phi_{ij}^\textrm{N}$ follows a Gaussian distribution with standard deviation, $\sigma_{\phi_{ij}^\textrm{N}} = (\sqrt{2}\,\rho_{ij}^\textrm{N})^{-1}$, when $\rho_{ij}^\textrm{N}\gg 1$ \cite{cra89}. From Eq.~(\ref{eqn:cpphase-sum}), for $\rho_{ij}^\textrm{N}\gg 1$, $\phi_\nabla^\textrm{m}$ will also approach a Gaussian distribution with variance, $\sigma_{\phi_\nabla^\textrm{m}}^2 = \sum_{ij}\,(\sqrt{2}\,\rho_{ij})^{-2}$, where, $\rho_{ij}=\rho_{ij}^\textrm{N}$ (if $\rho_{ij}^\textrm{N} < \rho_{ij}^\textrm{H{\sc i}}$) or $\rho_{ij}=\rho_{ij}^\textrm{H{\sc i}}$ (if $\rho_{ij}^\textrm{H{\sc i}} < \rho_{ij}^\textrm{N}$). $\delta\phi_\nabla^\textrm{H{\sc i}}$ can be made the dominant fluctuation in $\phi_\nabla^\textrm{m}$ by averaging a number of independent samples ($N_\textrm{m}$) of $\phi_\nabla^\textrm{m}$ and reducing the variance of $\delta\phi_\nabla^\textrm{N}$ to $\sigma_{\delta\phi_\nabla^\textrm{N}}^2 / N_\textrm{m}$.

\section{Modeling}\label{sec:modeling}

\subsection{Instrument Model}\label{sec:instrument}

We use the Hydrogen Epoch of Reionization Array \cite[HERA;][]{deb17,thy16,ewa16,neb16,patra17} as an example to demonstrate our approach. Located at the Karoo desert in South Africa at a latitude of $-30.72^\circ$, HERA will consist of 350 close-packed 14~m dishes with a shortest antenna spacing of 14.6~m. Its triple split-core hexagonal layout is optimized for both redundant calibration and the delay spectrum technique for detecting the cosmic EoR \HI\ signal \cite{dil16}.

\subsection{Sky Model and Noise}\label{sec:skymodel-noise}

We construct an all-sky model that includes the realization of a fiducial {\sc faint galaxies} EoR model \cite{gre17b} using 21cmFAST \cite{mes11} and a radio foreground model that includes compact and diffuse synchrotron emission from the Galaxy and extragalactic sources as in \cite{thy15a}.

Fig.~\ref{fig:vis-spectra} shows $|V_{ij}(f)|$ measured on three non-redundant 14.6~m antenna spacings due to the sky (foreground synchrotron and EoR \HI) and thermal noise contributions with 1~min integration. The foreground contributions exceed the EoR signal by a factor $\sim 10^4$ as expected. It is also noted that $\rho_{ij}^\textrm{N} \gtrsim 100$, which implies $\delta\phi_\nabla^\textrm{N}$ will follow a Gaussian distribution with variance, $\sigma_{\delta\phi_\nabla^\textrm{N}}^2$.

\begin{figure}[htb]
\includegraphics[width=\linewidth]{visamp_spectra_asm_eor_noise}
\caption{Spatial coherence amplitude spectra on \replaced[id=NT]{antenna pairs forming a 14.6~m equilateral triad}{three non-redundant 14.6~m antenna spacings} are shown in red, blue and cyan. Solid \added[id=NT]{and dotted} lines show amplitudes from synchrotron foregrounds \added[id=NT]{and a fiducial EoR model, respectively}. Gray curve shows the typical noise obtained with 1~min integration. \deleted[id=NT]{Dotted curves show spatial coherence amplitudes from a fiducial EoR signal.} Typically, the foregrounds are $\gtrsim 10^4$ times brighter than the EoR signal. The orange bands denote the frequency sub-bands centered on 125~MHz, 150~MHz, and 175~MHz each of 10~MHz effective bandwidth. \label{fig:vis-spectra}}
\end{figure}

Fig.~\ref{fig:cp-spectra} shows the the bi-spectrum phase spectra for the foreground and EoR components separately. Clearly, the foreground component is characterized by a smoother spectrum whereas the EoR component fluctuates rapidly. The observed $\phi_\nabla^\textrm{m}$, that includes the foregrounds, EoR \HI\, and noise components, will be described by Eq.~\ref{eqn:cpphase-sum}.

\begin{figure}[htb]
\includegraphics[width=\linewidth]{closure_phase_spectra_0_1_8}
\caption{Bi-spectrum phase spectra of individual components in the sky model -- single point source (black) with $\phi_\nabla=0$, compact sources (cyan), diffuse foregrounds (blue), diffuse and compact components combined (red), and the EoR \HI\ fluctuations (gray) -- for a 14.6~m equilateral triad. The \HI\ component is highly fluctuating relative to the foregrounds. The phase wrap discontinuities at $\phi_\nabla=\pm \pi$ are not of a physical origin. The orange sub-bands are same as in Fig.~\ref{fig:vis-spectra}. \label{fig:cp-spectra}}
\end{figure}

\section{Extraction of the cosmic signal}\label{sec:extraction}

The contrasting spectral characteristics -- smooth $\phi_\nabla^\textrm{F}$ and fluctuating $\delta\phi_\nabla^\textrm{H{\sc i}}$ -- indicate that techniques similar to the power spectrum approaches could be employed to separate the cosmic signal from foregrounds but avoiding the need for high-precision spectral calibration that most existing approaches rely on.

If a number of independent measurements are available, they could be used to average $\phi_\nabla^\textrm{m}$ to reduce the standard deviation of $\delta\phi_\nabla^\textrm{N}$ by taking advantage of its Gaussian distribution. When $\delta\phi_\nabla^\textrm{N}$ is sufficiently low, $\delta\phi_\nabla^\textrm{H{\sc i}}$ will dominate the spectral fluctuations in $\phi_\nabla^\textrm{m}$, while $\phi_\nabla^\textrm{F}$ will be dominant overall but spectrally smooth. The variance in $\delta\phi_\nabla^\textrm{H{\sc i}}$ is effectively a measure of the EoR \HI\ line strength to foreground continuum ratio.

Thus, when $\rho_{ij}^\textrm{H{\sc i}} < \rho_{ij}^\textrm{N}$, $\sigma_{\phi_\nabla^\textrm{m}}$ can be used to estimate $\rho_{ij}^\textrm{H{\sc i}}$, which in turn can be used to infer $|V_{ij}^\textrm{H{\sc i}}|$ if \replaced[id=NT]{$|V_{ij}^\textrm{F}|$ }{$V_{ij}^\textrm{F}$} is known. $|V_{ij}^\textrm{H{\sc i}}|$ is directly related to the EoR H\,{\sc i} brightness temperature fluctuations, $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}})^2\rangle$, which can be inferred as a function of  redshift. This is discussed further below.

\subsection{Delay Power Spectrum of Bi-spectrum Phase}\label{sec:cp-FG-wedge}

While any method that relies on separating a signal from contaminants using differences in spectral characteristics is applicable, we employ the {\it Delay Spectrum} approach \cite{par12a,par12b}. Since the $\phi_\nabla$ spectrum can have discontinuities at $\phi_\nabla=\pm\pi$, and hence non-physical spectral features, we define $\xi_\nabla = e^{i\phi_\nabla^\textrm{m}}$, which is unaffected by these discontinuities. We perform a delay transform:
\begin{align}\label{eqn:cpdspec}
  \Xi_\nabla(\tau) &= \int \xi_\nabla(f)\,W(f)\,e^{i2\pi f\tau}\,\mathrm{d}f,
\end{align}
where, $W(f)$ is a spectral weighting usually chosen to control the quality of the delay spectrum \citep{thy13,thy16} and has an effective bandwidth, $B_\textrm{eff}$. $\Xi_\nabla(\tau)$ has units of Hz.

We define the delay cross-power spectrum of $\xi_\nabla(f)$ as:
\begin{align}
  P_\nabla(k_\parallel) &\equiv \mathfrak{Re}\bigg\{\Xi_\nabla(\tau)\,\Xi_{\nabla^\prime}^*(\tau)\bigg\} \left(\frac{\Delta D}{B_\textrm{eff}^2}\right), \label{eqn:delay-power-spectrum}
\end{align}
where, $\Delta D\equiv \Delta D(z)$ is the comoving depth along the line of sight corresponding to $B_\textrm{eff}$ at redshift $z$, $k_\parallel\approx 2\pi\tau B_\textrm{eff}/\Delta D$ is line-of-sight wavenumber, and $\mathfrak{Re}\{\cdot\}$ denotes the real part. In this paper, we use cosmological parameters from \cite{planck15xiii} with $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$. $P_\nabla(k_\parallel)$ is in units of $h^{-1}$~Mpc. When $\Xi_\nabla(\tau)=\Xi_{\nabla^\prime}(\tau)$, $P_\nabla(k_\parallel)$ reduces to delay auto-power spectrum. In \S\ref{sec:averaging}, we describe the usage of Eq.~(\ref{eqn:delay-power-spectrum}) in different contexts.

From Eq.~(\ref{eqn:cpphase-sum}), since $\xi_\nabla = \prod_{ij} e^{i\phi_{ij}^\textrm{m}}$, it can be shown that $\Xi_\nabla(\tau) = \Xi_{ab}(\tau)\circledast \Xi_{bc}(\tau)\circledast \Xi_{ca}(\tau)\circledast \mathcal{W}(\tau)$, where, $\{\xi_{ij}(f),\,\Xi_{ij}(\tau)\}$, and $\{W(f),\,\mathcal{W}(\tau)\}$ are Fourier transform pairs, and $\circledast$ denotes convolution. Foreground contribution to spatial coherence is known to be confined to delay modes within the {\it horizon limits} \cite{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,thy15a,thy15b,thy16,dil14}. As a result of the convolution, the foreground contribution to $\Xi_\nabla(\tau)$ is expected to be more extended along $\tau$ (and $k_\parallel$) relative to $\Xi_{ij}(\tau)$. This relation will be explored in a forthcoming study.

\subsection{Improving sensitivity}\label{sec:averaging}

We require $\sigma_{\delta\phi_\nabla^\textrm{N}} < \sigma_{\delta\phi_\nabla^\textrm{H{\sc i}}}$ such that $\phi_\nabla^\textrm{F}, \delta\phi_\nabla^\textrm{H{\sc i}} > \delta\phi_\nabla^\textrm{N}$ in order to detect EoR and estimate $\rho_{ab}^\textrm{H{\sc i}}$. This can be achieved by a combination of the following.

First, $\xi_\nabla$ can be averaged coherently until the variation due to the transiting sky exceeds the decrease in noise due to averaging \cite{car18}. An Allan Variance estimate for HERA showed this timescale to be $\lesssim 2$~min \cite{car18}.

Second, $\xi_\nabla$ can be averaged coherently across multiple days at a fixed LST before computing $P_\nabla(k_\parallel)$, as long as the array size is small compared to the predominant spatial scales in the ionospheric variations, as has been observed to be the case for HERA \cite{car18}.

Third, the measurements on redundant triads can be averaged coherently in $\xi_\nabla$, before computing the power spectrum.  For non-redundant triads (including different triad classes -- equilateral, isosceles, and scalene), it may be possible to average incoherently in $P_\nabla(k_\parallel)$, using  $\Xi_\nabla(\tau)$ and $\Xi_{\nabla^\prime}^*(\tau)$ from triad pairs $\nabla$ and $\nabla^\prime$ respectively to further improve sensitivity.

Lastly, since the EoR signal is statistically isotropic in space while the foregrounds are not, measurements on time intervals larger than the coherence timescale can be used to compute the individual delay cross-power spectra, $P_\nabla(k_\parallel)$, from temporal pairs $\Xi_\nabla(\tau)$ and $\Xi_{\nabla^\prime}^*(\tau)$, and then be averaged incoherently.

\subsection{Detection of Cosmic Reionization}\label{sec:EoR-detection}

The primary application of this technique is to detect the EoR \HI\ signal. Using an {\it Airy} power pattern for the HERA dish, we used PRISim \footnote{The Precision Radio Interferometry Simulator (PRISim) Python package is available at \href{https://github.com/nithyanandan/PRISim}{https://github.com/nithyanandan/PRISim}.} to simulate $V_{ij}^\textrm{F}(f)$, $V_{ij}^\textrm{H{\sc i}}(f)$, and $V_{ij}^\textrm{N}(f)$ from the synchrotron foregrounds, an EoR \HI\ model, and noise, respectively as described in \S\ref{sec:modeling} over the 100--200~MHz band with 195.3125~kHz spectral resolution, and a conservative phase-coherent integration interval of 1~min \cite{car18}.

For bi-spectrum phase, we consider 14.6~m equilateral antenna triads and assume ideal redundancy for the array. The total number of such measurements is assumed to be $N_\textrm{m} \sim 10^6$ (after allowing for $\sim$~50\% efficiency of data quality from $\sim 30$ redundant 14.6~m equilateral triads with HERA-47 \added[id=NT]{\cite[see layout in][]{car18}}, $\sim$~8~hours of observing per day at 1~min integration intervals, repeated over $\sim$~150 nights). We assume there are $N_\textrm{c}$ coherent measurements of $\phi_\nabla^\textrm{m}$ and for each of these measurements, there are $N_\textrm{ic}$ incoherent measurements such that $N_\textrm{c}N_\textrm{ic}=N_\textrm{m}$. $\phi_\nabla^\textrm{m}$ is averaged coherently over $N_\textrm{c}$ measurements, and then averaged incoherently in $P_\nabla(k_\parallel)$ measured for all incoherent pairs of $\phi_\nabla^\textrm{m}$. We choose $W(f)$ to be the inverse Fourier transform of the squared delay response of a {\it Blackman-Harris} spectral weighting \cite{har78} as proposed in \cite{thy16}, on sub-bands centered at 125~MHz, 150~MHz, and 175~MHz, with an effective bandwidth of $B_\textrm{eff}\simeq 10$~MHz (to minimize EoR signal evolution over the redshift range).

Fig.~\ref{fig:cpdps} shows the delay cross-power spectra, $P_\nabla(k_\parallel)$, of $\phi_\nabla^\textrm{m}$ obtained for the three sub-bands, with an initial S/N, $\rho_{ij}^\textrm{N}=200$. It is seen that the contributions from the EoR \HI\ fluctuations, $\delta\phi_\nabla^\textrm{H{\sc i}}$, are detected as they dominate over $\phi_\nabla^\textrm{F} + \delta\phi_\nabla^\textrm{N}$ at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$ in the 150~MHz and 175~MHz sub-bands. The absence of a clear detection in the 125~MHz sub-band is discussed below.

\begin{figure*}[htb]
  \includegraphics[width=\linewidth]{cpdps_all_subbands_SNR_200_nsamples_1048576}
\caption{Delay power spectra of bi-spectrum phases in sub-bands centered at 125~MHz (left), 150~MHz (middle), and 175~MHz (right) that include contributions from foregrounds, EoR signal, and noise. \added[id=NT]{The corresponding redshifts are listed in each panel. The 14.6~m equilateral triad used is highlighted on a HERA-19 layout on the top left.} \replaced[id=NT]{Green curve shows only the foreground contribution ($\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}$).}{The orange curve applies when only foregrounds are present, $\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}$.} The anisotropic foreground model employed \cite{thy15a} results in the asymmetry around $k_\parallel=0\,h^{-1}$~Mpc. The red curve \replaced[id=NT]{shows the effect of including}{is when} noise fluctuations \replaced[id=NT]{($\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}+\delta\phi_\nabla^\textrm{N}$)}{for noise fluctuations, $\delta\phi_\nabla^\textrm{N}$ superimposed on the foreground component, $\phi_\nabla^\textrm{F}$}, after averaging $N_\textrm{m} \sim 10^6$ measurements. The gray curve is when fluctuations from the EoR \HI\ signal are \replaced[id=NT]{included}{also present}, $\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}+\delta\phi_\nabla^\textrm{H{\sc i}}+\delta\phi_\nabla^\textrm{N}$. The black curve is \replaced[id=NT]{identical to}{the same as} the gray one except the EoR \HI\ signal is 10 times brighter. The dot and vertical markers denote positive and negative values respectively. The latter results from the removal of noise bias by using cross-power spectra (Eq.~\ref{eqn:delay-power-spectrum}). The EoR \HI\ contribution is significantly detected at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$ at $z\approx 8.5$ and $z\approx 7.1$. At $z\approx 10.4$, the EoR contribution is \deleted[id=NT]{virtually} inseparable from \deleted[id=NT]{the} foregrounds because \replaced[id=NT]{the EoR signal is weaker (see power spectrum of {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b}) while the foregrounds are brighter}{they are brighter and the EoR signal is weaker (see power spectrum of {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b})}. The detection significance depends on the strength of EoR H\,{\sc i} signal relative to the foregrounds. A 10-fold increase in EoR intensity shows a $\sim 100$-fold increase in $P_\nabla(k_\parallel)$ at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$ at $z\approx 8.5$ and $z\approx 7.1$, indicating its sensitive dependence on $\rho_{ij}^\textrm{H{\sc i}}$. This detection measures the line-to-continuum ratio, from which the EoR H\,{\sc i} brightness temperature fluctuations, $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$, can be inferred using a reliable foreground model. \label{fig:cpdps}}
\end{figure*}

\subsection{EoR H\,{\sc i} Brightness Temperature Fluctuations} \label{sec:Tb-variance}

Beyond simple detection, we could also estimate the EoR \HI\ brightness temperature fluctuations, $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$, at various redshifts using the separation in $P_\nabla(k_\parallel)$. When $\rho_{ij}^\textrm{H{\sc i}} < \rho_{ij}^\textrm{N}$, the separation between $\phi_\nabla^\textrm{m}$ and $\phi_\nabla^\textrm{F} + \delta\phi_\nabla^\textrm{N}$ in $P_\nabla(k_\parallel)$ depends sensitively on $\rho_{ab}^\textrm{H{\sc i}}$. Fig.~\ref{fig:cpdps} shows $P_\nabla(k_\parallel)$ in the three sub-bands for different values of $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$, or equivalently, $\rho_{ij}^\textrm{H{\sc i}}$. As the EoR \HI\ signal strength increases (ten-fold in $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle^{1/2}$), the separation in $P_\nabla(k_\parallel)$ also increases at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$ ($\sim 100$ times), most clearly in the 150~MHz and 175~MHz sub-bands. Conversely, this separability can be used to estimate $\rho_{ij}^\textrm{H{\sc i}}$ in the different sub-bands, which in turn can be used to estimate $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$ by using the best foreground models available in those sub-bands. The accuracy of $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$ so determined will depend on the uncertainty in the foreground model, which is not required to be as precise as the calibration requirement.

The redshift evolution of separability in $P_\nabla(k_\parallel)$ depends on relative strengths of the foreground and the EoR \HI\ signal. In the 175~MHz sub-band ($z\approx 7.1$), relative to the 150~MHz sub-band ($z\approx 8.5$), $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$ is weaker (see {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b}) but the foreground synchrotron temperature is also fainter roughly by the same factor. Therefore, the separability of EoR \HI\ in $P_\nabla(k_\parallel)$ is roughly similar in the two sub-bands (Fig.~\ref{fig:cpdps-150}, \ref{fig:cpdps-175}). However, in the 125~MHz sub-band ($z\approx 10.4$), $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$ in the fiducial EoR model is fainter but the foregrounds are brighter, thereby significantly decreasing the spectral contrast between foregrounds and the EoR signal. Hence, a clear separability is not seen at $z\approx 10.4$ (Fig.~\ref{fig:cpdps-125}).

This hypothesis can be confirmed by estimating $\rho_{ij}^\textrm{H{\sc i}}$ in different sub-bands and verifying that there is no separation in $P_\nabla(k_\parallel)$ at any $k$ at $z\lesssim 6$, when the IGM is understood to be fully ionized, with a reasoning similar to \cite{pob16b}. Lack of separation in $P_\nabla(k_\parallel)$, such as at $z\approx 10.4$, will enable placing upper limits on $\langle (\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$.

\section{Summary}\label{sec:summary}

A primary limitation to existing low-frequency EoR 21~cm cosmology experiments remains high-precision spectral calibration in order to remove the strong continuum foregrounds. We propose a new approach that uses bi-spectrum phase, which represents an intrinsic property of the sky and is thus impervious to antenna-based calibration and associated errors, unlike existing approaches.

The primary goal of the proposed technique is to obtain an interferometric detection of the \HI\ 21~cm signal from the neutral IGM during cosmic reionization. The next goal will be to relate the measured EoR \HI\ bi-spectrum phase power spectrum to the astrophysical quantities of interest, particularly, the brightness temperature fluctuations. The magnitude of the line signal relative to that of the foreground continuum in the bi-spectrum phase power spectrum, is essentially a measure of the line-to-continuum ratio, since the line signal introduces small, frequency-dependent, perturbations on the spatial coherence phase. Hence, the line signal strength can be calibrated against a foreground model.

Ultimately, the quantitative interpretation will entail a standard forward-modeling process, comparing measurements to the predictions of bi-spectrum phase power spectra for different sky models. This work investigates one such realization, and demonstrates that a quantitative relationship exists between the EoR signal strength and the bi-spectrum phase power spectra.

Beyond the simple demonstration presented here, there are many ways to improve the prospects of detecting the cosmic reionization signal using the bi-spectrum phase as outlined in \S\ref{sec:averaging} using optimal weighting techniques \cite{liu14a,liu14b,dil15}, particularly in the context of a redundant array such as HERA (though it applies to non-redundant arrays as well). Upon completion, HERA will consist of 350 antennas yielding $\sim$~10--100 times more measurements than used here, and thereby go deeper in overall sensitivity. Removal of the foreground contribution in $\xi_\nabla$ could potentially relax the dynamic range required for detection and boost sensitivity in $|k_\parallel| \lesssim 0.5\,h$~Mpc$^{-1}$. 
% % Challenges to this approach arise from effects that introduce systematic spectral errors in Eq.~(\ref{eqn:cpphase-sum}). 
% % Due to the triple convolution in Eq.~(\ref{eqn:cpdspec-convolution}), there may be  fewer sensitive $k_\parallel$-modes. 
% Systematics due to the presence of significant multiplicative gain terms that cannot be factorized into antenna-dependent terms, polarization leakage between the dual-polarization feeds, and direction-dependent antenna gains may challenge this approach.
% % , all of which will introduce systematic departures from Eq.~(\ref{eqn:cpphase-sum}). 
% Excision of radio frequency interference (RFI) affected frequencies, if not carefully corrected for, will introduce spectral artifacts, and reduce the sensitivity in relatively foreground-free $k_\parallel$-modes. However, it must be noted that most existing approaches face these challenges as well. 
% In future work, we will investigate the challenges and the detectability of EoR using this approach on HERA data.

\begin{acknowledgments}
We acknowledge the insightful and helpful inputs from Gianni Bernardi, Judd Bowman, James Aguirre, Aaron Parsons, Jonathan Pober, Adam Beardsley, and Peter Williams that helped in the preparation of this manuscript. The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc. We acknowledge support from the Royal Society and the Newton Fund under grant NA150184.
\end{acknowledgments}

\bibliography{eor}

\end{document}

### 2018-04-27 ms_changes.tex
Update on Overleaf.

@@
 \usepackage[caption=false]{subfig}
-\usepackage[final]{changes}
+\usepackage[draft]{changes}
 \definechangesauthor[name={Nithyanandan Thyagarajan}, color=red]{NT}
@@
 \begin{acknowledgments}
-We acknowledge the insightful and helpful inputs from Gianni Bernardi, Judd Bowman, James Aguirre, Aaron Parsons, Jonathan Pober, Adam Beardsley, and Peter Williams that helped in the preparation of this manuscript.
+We acknowledge the insightful and helpful inputs from Rajaram Nityananda, James Moran, Gianni Bernardi, Judd Bowman, James Aguirre, Aaron Parsons, Jonathan Pober, Adam Beardsley, and Peter Williams that helped in the preparation of this manuscript.
 The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.

### 2018-04-27 RefereeReportResponse-17Apr2018.txt
Update on Overleaf.

@@
+We thank the referees for their valuable comments and suggestions. We
+are very delighted that both reviews have been very positive. In 
+addressing the comments, we find that the manuscript has gained in
+quality. We have responded to specific comments (see below).
+
 ----------------------------------------------------------------------
@@
 Introduction about applicability to other 21 cm intensity mapping
-experiments that will measure the BAO.
+experiments that will measure the BAO, for example.
 

### 2018-04-27 ResponseToEditor.txt
Update on Overleaf.

To the editorial staff:

We appreciate the referees' reports and their positive views on 
our work. We have responded to their comments, attached in a 
separate file.

We are attaching two versions of the Latex file. One is the 
updated ms.tex (ms.pdf). The other is ms_changes.tex (and 
ms_changes.pdf) and it contain all the changes highlighted in 
red.

Regarding the length of the manuscript, we believe there has 
been a mistake. In the email from PRL on 7 March 2018, the word 
count summary was stated to be:

-----------------------------------------------------------------
  Figure   Aspect Ratio   Wide?   Word Equivalent
     1         1.54         No        117
     2         1.06         No        161
     3         3.73        Yes        200

WORD COUNT SUMMARY
    Note: Text word count excludes title, abstract, byline, PACS,
          receipt date, acknowledgments, and references.

              Text word count   4003
        Equations word equiv.    192
          Figures word equiv.    478
           Tables word equiv.      0
                              ------
                       TOTAL    4673 (Maximum length is 3750)
-----------------------------------------------------------------

Subsequently, we made substantial changes to reduce the length by
at least 100 words including substantial changes to figure sizes. 
We confirmed this through "texcount" on the command line and 
carefully kept track of the word count at the time of initial 
submission (5 March 2018) as well as the trimmed one (12 March 
2018).

In the recent email from PRL on 17 April 2018, the word count 
was found by PRL to be: 
-----------------------------------------------------------------
  Figure   Aspect Ratio   Wide?   Word Equivalent
     1         1.54         No        117
     2         1.06         No        161
     3         3.73        Yes        200

WORD COUNT SUMMARY
    Note: Text word count excludes title, abstract, byline, PACS,
          receipt date, acknowledgments, and references.

              Text word count   4003
        Equations word equiv.    192
          Figures word equiv.    478
           Tables word equiv.      0
                              ------
                       TOTAL    4516 (Maximum length is 3750)
-----------------------------------------------------------------
You will notice that the word count for the text, equations, and 
figures are exactly identical between 17 April and 7 March (which
is impossible), although the total has changed from 4673 to 4516. 
Not only has the itemized count not been updated, but the new 
total is also inaccurate.

We understand the output of texcount may not be identical to the 
software PRL uses. Therefore, we request PRL to give us an 
accurate count which we can use as a reference to further trim 
the manuscript, if necessary.

In the meanwhile, we request that the editorial office pass on
the revised manuscript to the referees for a speedy acceptance.

Best regards,
The authors.

### 2018-04-27 ms_changes.tex
Update on Overleaf.

--- 2018-04-26 ms_changes.tex
@@
 \usepackage[caption=false]{subfig}
-\usepackage[final]{changes}
-\definechangesauthor[name={Nithyanandan Thyagarajan}, color=red]{NT}
-\definechangesauthor[name={Chris Carilli}, color=blue]{CC}
-\definechangesauthor[name={Bojan Nikolic}, color=green]{BN}
-\setremarkmarkup{(#2)}
+\usepackage[draft]{changes}
+% \definechangesauthor[name={Nithyanandan Thyagarajan}, color=red]{NT}
+% \setremarkmarkup{(#2)}
+\colorlet{Changes@Color}{red}
+
 % \usepackage[mathlines]{lineno}% Enable numbering of text and display math
@@
 Current techniques require precise instrumental calibration to separate the weak H\,{\sc i} line signal from the foreground continuum emission.
-We propose to mitigate this calibration requirement by using \replaced[id=NT]{measurements of the interferometric bi-spectrum phase}{the phase of the bi-spectrum from interferometric measurements}.
+We propose to mitigate this calibration requirement by using \replaced{measurements of the interferometric bi-spectrum phase}{the phase of the bi-spectrum from interferometric measurements}.
 Bi-spectrum phase is unaffected by antenna-based direction-independent calibration errors and hence for a compact array it depends on the sky brightness distribution only (subject to the usual thermal-like noise).
@@
 Using fiducial models for continuum foregrounds, and for the cosmological H\,{\sc i} signal, we show the latter should be detectable in bi-spectrum phase spectra, with reasonable significance at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$, using existing instruments.
-\added[id=NT]{Our approach will also benefit other H\,{\sc i} intensity mapping experiments that face similar challenges, such as those measuring Baryon Acoustic Oscillations (BAO).}
+\added{Our approach will also benefit other H\,{\sc i} intensity mapping experiments that face similar challenges, such as those measuring Baryon Acoustic Oscillations (BAO).}
 \end{abstract}
@@
 Current indirect constraints suggest this second transition occurred a few hundred Myr to $\sim 1$~Gyr ($z\gtrsim 6$) \cite{gre17a} after the Big Bang.
-During this epoch, the photon-to-baryon ratio \replaced[id=NT]{rose significantly}{eventually exceeded unity}, reionizing the entire Universe except gas bound to galaxies.
+During this epoch, the photon-to-baryon ratio \replaced{rose significantly}{eventually exceeded unity}, reionizing the entire Universe except gas bound to galaxies.
 The astrophysical processes during this epoch shaped the evolution of galaxies and large-scale structure.
@@
 \cite{par10,tin13,van13}), with sensitivities sufficient for a statistical detection of the EoR \HI\ signal using power spectrum methodology \cite{thy13,bea13}.
-However, despite obtaining \replaced[id=NT]{significant}{sufficient} observational data, these experiments remain dynamic range limited due to the coupling of the strong foreground emission to the EoR \HI\ signal via instrumental effects, particularly systematics related to precision of calibration, and the intrinsic chromatic instrumental response \cite{thy15a,thy15b,thy16,dat09,dat10,tro16, pac13,ali15,patil17,pob15,liu10,zhe14,barry16,sie17,dil17}.
+However, despite obtaining \replaced{significant}{sufficient} observational data, these experiments remain dynamic range limited due to the coupling of the strong foreground emission to the EoR \HI\ signal via instrumental effects, particularly systematics related to precision of calibration, and the intrinsic chromatic instrumental response \cite{thy15a,thy15b,thy16,dat09,dat10,tro16, pac13,ali15,patil17,pob15,liu10,zhe14,barry16,sie17,dil17}.
 
@@
 
-\added[id=NT]{Further, \HI\ intensity mapping experiments, such as CHIME \cite{ban14}, HIRAX \cite{new16}, and Tianlai \cite{xu15} will map \HI\ on large-scales at $1\lesssim z \lesssim 3$ tracing the matter density fluctuations, and thus measure the Baryon Acoustic Oscillations (BAO) that can be used as a standard ruler in constraining the {\it Dark Energy} equation of state.
+\added{Further, \HI\ intensity mapping experiments, such as CHIME \cite{ban14}, HIRAX \cite{new16}, and Tianlai \cite{xu15} will map \HI\ on large-scales at $1\lesssim z \lesssim 3$ tracing the matter density fluctuations, and thus measure the Baryon Acoustic Oscillations (BAO) that can be used as a standard ruler in constraining the {\it Dark Energy} equation of state.
 The challenges in these experiments are expected to be of a similar magnitude to that in 21~cm EoR experiments, and thus can potentially benefit from our approach.}
@@
 An interferometer measures $V_{ij}^\textrm{m}(f) = g_i(f)\, g_j^*(f)\, V_{ij}^\textrm{T}(f) + V_{ij}^\textrm{N}(f)$, the  true spatial coherence, $V_{ij}^\textrm{T}(f)$, corrupted by multiplicative antenna gains and ionospheric distortions, $g_i$, and additive thermal-like noise, $V_{ij}^\textrm{N}(f)$.
-\replaced[id=NT]{We}{It is pertinent to} express $V_{ij}^\textrm{T}(f) = V_{ij}^\textrm{F}(f) + V_{ij}^\textrm{H{\sc i}}(f)$, where, $V_{ij}^\textrm{F}(f)$ and $V_{ij}^\textrm{H{\sc i}}(f)$ are \replaced[id=NT]{due to}{contributions from} foreground synchrotron and EoR \HI\ signal, respectively.
+\replaced{We}{It is pertinent to} express $V_{ij}^\textrm{T}(f) = V_{ij}^\textrm{F}(f) + V_{ij}^\textrm{H{\sc i}}(f)$, where, $V_{ij}^\textrm{F}(f)$ and $V_{ij}^\textrm{H{\sc i}}(f)$ are \replaced{due to}{contributions from} foreground synchrotron and EoR \HI\ signal, respectively.
 The measured bi-spectrum is $B_\nabla^\textrm{m}=\prod_{ij} V_{ij}^\textrm{m}=B_\nabla^\textrm{T}\prod_i |g_i|^2+B_\nabla^\textrm{N}$, where, the dependence on $f$ has been omitted for convenience (hereafter so, unless indicated).
-\added[id=NT]{In our notation, $ij$ only indexes three specific cyclic antenna pairs $\{ab, bc, ca\}$ and not all possible combinations of $i$ and $j$.} \deleted[id=NT]{The noise term,} $B_\nabla^\textrm{N}$, \added[id=NT]{encompassing all terms containing $V_{ij}^\textrm{N}$ including cross-terms,} has zero mean since it contains $V_{ij}^\textrm{N}$, \added[id=NT]{which is assumed to be uncorrelated between antenna pairs}.
+\added{In our notation, $ij$ only indexes three specific cyclic antenna pairs $\{ab, bc, ca\}$ and not all possible combinations of $i$ and $j$.} \deleted{The noise term,} $B_\nabla^\textrm{N}$, \added{encompassing all terms containing $V_{ij}^\textrm{N}$ including cross-terms,} has zero mean since it contains $V_{ij}^\textrm{N}$, \added{which is assumed to be uncorrelated between antenna pairs}.
 The phase of $B_\nabla^\textrm{m}$ is:
@@
 where, $\phi_{ij}^\textrm{m} = \phi_i - \phi_j + \phi_{ij}^\textrm{T} + \delta\phi_{ij}^\textrm{N}$, and $\phi_i$ are the phases of $V_{ij}^\textrm{m}$ and $g_i$, respectively.
-$\delta\phi_\nabla^\textrm{H{\sc i}}$ and $\delta\phi_\nabla^\textrm{N}$ \added[id=NT]{are not just determined by the phases of $V_{ij}^\textrm{H{\sc i}}$ and $V_{ij}^\textrm{N}$ but are small} \deleted[id=NT]{may be considered as small} perturbations due to \replaced[id=NT]{the EoR \HI\ signal and noise}{noise and EoR \HI\ signal} respectively, superimposed on dominant phase, $\phi_\nabla^\textrm{F}$, due to foregrounds, \added[id=NT]{and depend on $\rho_{ij}^\textrm{H{\sc i}}\equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{H{\sc i}}|$ and $\rho_{ij}^\textrm{N}\equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{N}|$ respectively.} \replaced[id=NT]{The small-perturbation formalism}{and} is valid when \replaced[id=NT]{$\rho_{ij}^\textrm{H{\sc i}}\gg 1$ and $\rho_{ij}^\textrm{N}\gg 1$}{$\rho_{ij}^\textrm{N} \equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{N}| \gg 1$ and $\rho_{ij}^\textrm{H{\sc i}} \equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{H{\sc i}}| \gg 1$}, as is usually the case in \replaced[id=NT]{low-frequency cosmology}{EoR \HI\ } experiments.
+$\delta\phi_\nabla^\textrm{H{\sc i}}$ and $\delta\phi_\nabla^\textrm{N}$ \added{are not just determined by the phases of $V_{ij}^\textrm{H{\sc i}}$ and $V_{ij}^\textrm{N}$ but are small} \deleted{may be considered as small} perturbations due to \replaced{the EoR \HI\ signal and noise}{noise and EoR \HI\ signal} respectively, superimposed on dominant phase, $\phi_\nabla^\textrm{F}$, due to foregrounds, \added{and depend on $\rho_{ij}^\textrm{H{\sc i}}\equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{H{\sc i}}|$ and $\rho_{ij}^\textrm{N}\equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{N}|$ respectively.} \replaced{The small-perturbation formalism}{and} is valid when \replaced{$\rho_{ij}^\textrm{H{\sc i}}\gg 1$ and $\rho_{ij}^\textrm{N}\gg 1$}{$\rho_{ij}^\textrm{N} \equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{N}| \gg 1$ and $\rho_{ij}^\textrm{H{\sc i}} \equiv |V_{ij}^\textrm{F}|/|V_{ij}^\textrm{H{\sc i}}| \gg 1$}, as is usually the case in \replaced{low-frequency cosmology}{EoR \HI\ } experiments.
 $\phi_\nabla^\textrm{m}$ is independent of the antenna gains and measures $\phi_\nabla^\textrm{T}$, corrupted only by noise, $\delta\phi_\nabla^\textrm{N}$.
@@
 \includegraphics[width=\linewidth]{visamp_spectra_asm_eor_noise}
-\caption{Spatial coherence amplitude spectra on \replaced[id=NT]{antenna pairs forming a 14.6~m equilateral triad}{three non-redundant 14.6~m antenna spacings} are shown in red, blue and cyan.
-Solid \added[id=NT]{and dotted} lines show amplitudes from synchrotron foregrounds \added[id=NT]{and a fiducial EoR model, respectively}.
+\caption{Spatial coherence amplitude spectra on \replaced{antenna pairs forming a 14.6~m equilateral triad}{three non-redundant 14.6~m antenna spacings} are shown in red, blue and cyan.
+Solid \added{and dotted} lines show amplitudes from synchrotron foregrounds \added{and a fiducial EoR model, respectively}.
 Gray curve shows the typical noise obtained with 1~min integration.
-\deleted[id=NT]{Dotted curves show spatial coherence amplitudes from a fiducial EoR signal.} Typically, the foregrounds are $\gtrsim 10^4$ times brighter than the EoR signal.
+\deleted{Dotted curves show spatial coherence amplitudes from a fiducial EoR signal.} Typically, the foregrounds are $\gtrsim 10^4$ times brighter than the EoR signal.
 The orange bands denote the frequency sub-bands centered on 125~MHz, 150~MHz, and 175~MHz each of 10~MHz effective bandwidth.
@@
 
-Thus, when $\rho_{ij}^\textrm{H{\sc i}} < \rho_{ij}^\textrm{N}$, $\sigma_{\phi_\nabla^\textrm{m}}$ can be used to estimate $\rho_{ij}^\textrm{H{\sc i}}$, which in turn can be used to infer $|V_{ij}^\textrm{H{\sc i}}|$ if \replaced[id=NT]{$|V_{ij}^\textrm{F}|$ }{$V_{ij}^\textrm{F}$} is known.
+Thus, when $\rho_{ij}^\textrm{H{\sc i}} < \rho_{ij}^\textrm{N}$, $\sigma_{\phi_\nabla^\textrm{m}}$ can be used to estimate $\rho_{ij}^\textrm{H{\sc i}}$, which in turn can be used to infer $|V_{ij}^\textrm{H{\sc i}}|$ if \replaced{$|V_{ij}^\textrm{F}|$ }{$V_{ij}^\textrm{F}$} is known.
 $|V_{ij}^\textrm{H{\sc i}}|$ is directly related to the EoR H\,{\sc i} brightness temperature fluctuations, $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}})^2\rangle$, which can be inferred as a function of  redshift.
@@
 For bi-spectrum phase, we consider 14.6~m equilateral antenna triads and assume ideal redundancy for the array.
-The total number of such measurements is assumed to be $N_\textrm{m} \sim 10^6$ (after allowing for $\sim$~50\% efficiency of data quality from $\sim 30$ redundant 14.6~m equilateral triads with HERA-47 \added[id=NT]{\cite[see layout in][]{car18}}, $\sim$~8~hours of observing per day at 1~min integration intervals, repeated over $\sim$~150 nights).
+The total number of such measurements is assumed to be $N_\textrm{m} \sim 10^6$ (after allowing for $\sim$~50\% efficiency of data quality from $\sim 30$ redundant 14.6~m equilateral triads with HERA-47 \added{\cite[see layout in][]{car18}}, $\sim$~8~hours of observing per day at 1~min integration intervals, repeated over $\sim$~150 nights).
 We assume there are $N_\textrm{c}$ coherent measurements of $\phi_\nabla^\textrm{m}$ and for each of these measurements, there are $N_\textrm{ic}$ incoherent measurements such that $N_\textrm{c}N_\textrm{ic}=N_\textrm{m}$.
@@
 \caption{Delay power spectra of bi-spectrum phases in sub-bands centered at 125~MHz (left), 150~MHz (middle), and 175~MHz (right) that include contributions from foregrounds, EoR signal, and noise.
-\added[id=NT]{The corresponding redshifts are listed in each panel.
-The 14.6~m equilateral triad used is highlighted on a HERA-19 layout on the top left.} \replaced[id=NT]{Green curve shows only the foreground contribution ($\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}$).}{The orange curve applies when only foregrounds are present, $\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}$.} The anisotropic foreground model employed \cite{thy15a} results in the asymmetry around $k_\parallel=0\,h^{-1}$~Mpc.
-The red curve \replaced[id=NT]{shows the effect of including}{is when} noise fluctuations \replaced[id=NT]{($\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}+\delta\phi_\nabla^\textrm{N}$)}{for noise fluctuations, $\delta\phi_\nabla^\textrm{N}$ superimposed on the foreground component, $\phi_\nabla^\textrm{F}$}, after averaging $N_\textrm{m} \sim 10^6$ measurements.
-The gray curve is when fluctuations from the EoR \HI\ signal are \replaced[id=NT]{included}{also present}, $\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}+\delta\phi_\nabla^\textrm{H{\sc i}}+\delta\phi_\nabla^\textrm{N}$.
-The black curve is \replaced[id=NT]{identical to}{the same as} the gray one except the EoR \HI\ signal is 10 times brighter.
+\added{The corresponding redshifts are listed in each panel.
+The 14.6~m equilateral triad used is highlighted on a HERA-19 layout on the top left.} \replaced{Green curve shows only the foreground contribution ($\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}$).}{The orange curve applies when only foregrounds are present, $\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}$.} The anisotropic foreground model employed \cite{thy15a} results in the asymmetry around $k_\parallel=0\,h^{-1}$~Mpc.
+The red curve \replaced{shows the effect of including}{is when} noise fluctuations \replaced{($\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}+\delta\phi_\nabla^\textrm{N}$)}{for noise fluctuations, $\delta\phi_\nabla^\textrm{N}$ superimposed on the foreground component, $\phi_\nabla^\textrm{F}$}, after averaging $N_\textrm{m} \sim 10^6$ measurements.
+The gray curve is when fluctuations from the EoR \HI\ signal are \replaced{included}{also present}, $\phi_\nabla^\textrm{m}=\phi_\nabla^\textrm{F}+\delta\phi_\nabla^\textrm{H{\sc i}}+\delta\phi_\nabla^\textrm{N}$.
+The black curve is \replaced{identical to}{the same as} the gray one except the EoR \HI\ signal is 10 times brighter.
 The dot and vertical markers denote positive and negative values respectively.
@@
 The EoR \HI\ contribution is significantly detected at $|k_\parallel| \gtrsim 0.5\,h$~Mpc$^{-1}$ at $z\approx 8.5$ and $z\approx 7.1$.
-At $z\approx 10.4$, the EoR contribution is \deleted[id=NT]{virtually} inseparable from \deleted[id=NT]{the} foregrounds because \replaced[id=NT]{the EoR signal is weaker (see power spectrum of {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b}) while the foregrounds are brighter}{they are brighter and the EoR signal is weaker (see power spectrum of {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b})}.
+At $z\approx 10.4$, the EoR contribution is \deleted{virtually} inseparable from \deleted{the} foregrounds because \replaced{the EoR signal is weaker (see power spectrum of {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b}) while the foregrounds are brighter}{they are brighter and the EoR signal is weaker (see power spectrum of {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b})}.
 The detection significance depends on the strength of EoR H\,{\sc i} signal relative to the foregrounds.

### 2018-04-27 ResponseToEditor.txt
PRL-26Apr2018-submitted

@@
 updated ms.tex (ms.pdf). The other is ms_changes.tex (and 
-ms_changes.pdf) and it contain all the changes highlighted in 
+ms_changes.pdf) and it contains all the changes highlighted in 
 red.
@@
 Subsequently, we made substantial changes to reduce the length by
-at least 100 words including substantial changes to figure sizes. 
+at least 1000 words including substantial changes to figure sizes. 
 We confirmed this through "texcount" on the command line and 

### 2018-04-27 ms_changes.tex
Update on Overleaf.

@@
 In the 175~MHz sub-band ($z\approx 7.1$), relative to the 150~MHz sub-band ($z\approx 8.5$), $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$ is weaker (see {\sc faint galaxies} model at $k=0.5$~Mpc$^{-1}$ in Fig.~2 of \cite{gre17b}) but the foreground synchrotron temperature is also fainter roughly by the same factor.
-Therefore, the separability of EoR \HI\ in $P_\nabla(k_\parallel)$ is roughly similar in the two sub-bands (Fig.~\ref{fig:cpdps-150}, \ref{fig:cpdps-175}).
+Therefore, the separability of EoR \HI\ in $P_\nabla(k_\parallel)$ is roughly similar in the two sub-bands.
 However, in the 125~MHz sub-band ($z\approx 10.4$), $\langle(\delta T_\textrm{b}^\textrm{H{\sc i}}(z))^2\rangle$ in the fiducial EoR model is fainter but the foregrounds are brighter, thereby significantly decreasing the spectral contrast between foregrounds and the EoR signal.
-Hence, a clear separability is not seen at $z\approx 10.4$ (Fig.~\ref{fig:cpdps-125}).
+Hence, a clear separability is not seen at $z\approx 10.4$.
 

### 2018-05-01 ms_changes.tex
Update Acknowledgements section (Add RN, JM)

@@
 \begin{acknowledgments}
-We acknowledge the insightful and helpful inputs from Gianni Bernardi, Judd Bowman, James Aguirre, Aaron Parsons, Jonathan Pober, Adam Beardsley, and Peter Williams that helped in the preparation of this manuscript.
+We acknowledge the insightful and helpful inputs from \added{Rajaram Nityananda}, Gianni Bernardi, Judd Bowman, James Aguirre, Aaron Parsons, \added{James Moran,} Jonathan Pober, Adam Beardsley, and Peter Williams that helped in the preparation of this manuscript.
 The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.
